---
title: "p8015_hw1_yl4924"
author: "Yongzheng Li"
date: "2022-09-24"
output: github_document
---
# Load the package
```{r}
library(tidyverse)
```

# Problem 1

## Load the 'penguin' dataset
```{r}
data("penguins", package = "palmerpenguins")
```

## Display the 'penguin' dataset
```{r}
head(penguins)
summary(penguins)
nrow(penguins)
ncol(penguins)
```


The dataset 'penguin' has 8 columns and 344 rows. The 8 columns are "species", "island", "bill_length_mm", "bill_depth_mm" , "flipper_length_mm", "body_mass_g" "sex" and "year". The mean bill_length, mean bill_depth, mean flipper_length, and mean body_mass for the dataset is 43.92mm, 17.15mm, 200.9mm, 4202g, respectively. It is worth to note that this dataset contain missing value.

## Make the scatter plot
```{r}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()
```

## Save the scatter plot in project directory
```{r}
ggsave("scatter_plot.pdf", height = 4, width = 6)
```

# Problem 2

## Create Data frame 
```{r}
problem_2_df = tibble(
  vec_random_sample = rnorm(10),
  vec_logic = vec_random_sample > 0,
  vec_char = c(rep("hw1",10)),
  vec_factor = factor(c(rep("H",3), 
                         rep("M", 4), 
                         rep("L", 3)), 
                       levels = c("L","M","H"))
)
```


## Take the mean of each variable
```{r}
mean(pull(problem_2_df, vec_random_sample))
mean(pull(problem_2_df, vec_logic))
```

After using mean function, random_sample and logic variable can return the result. Char variable and factor variable cannot return result.

## Convert char variable, logic variable, factor variable to numeric
```{r eval=FALSE}
as.numeric(pull(problem_2_df, vec_logic))
as.numeric(pull(problem_2_df, vec_char))
as.numeric(pull(problem_2_df, vec_factor))
```
 
For logic variable, the element become values of 1 and 0. 1 indicates the statement is TRUE, which the element in random sample > 0. 0 indicates the statement is FALSE, which the elements in random sample <= 0. This helps when I try to take the mean because the mean = 1*number of True element / total number of element.

For char variable, all the elements become NA and the console show warning that "argument is not numeric or logical: returning NA". It indicates character variable cannot be convert to numerical variable directly. It helps explain mean function does not work for this variable.

For factor variable, the result returns "3 3 3 2 2 2 2 1 1 1" in my cases. Since I assign the level for the factor variable, the result value correspond to the level I assign, "L" to 1, "M" to 2, "H" to 3. However, these values are not meaningful because "H" does not equal to 3 times "L". It helps explain why mean function does not works for this variable because the mean of factor variable is meaningless.

